\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2018-02-23 redaction/dhp_july2017.tex
changement texte

\begin{abstract}
This paper investigates the underlying sources behind the downward trend of international transport costs, exploiting information contained in the US imports flows database over 1974-2013 distinguished by transport mode, air and vessel. This drives us to provide a careful modeling of international transport costs, that allows for both ad-valorem (multiplicative) and per-unit (additive) components. In contrast to \cite{hummels2007}, we find that composition effects play a minor role in accounting for the reduction of transport costs relative to the value of US imports. Transport cost ``per se'' did decline. Importantly, this difference of results can be attributed to the new way of modeling the additive component of transport costs we offer. Further, in quantitative terms, we find that the additive component of transport costs is sizeable, representing between 40 and 50\% of overall transport costs. In both aspects, our results point the importance of the additive component in accounting for international transport costs.
\end{abstract}

@@

\textbf{j'ai enleve les references a Guillaume parce que ca me semble pas bien place ici, je ne sais pas bien ce que ca veut dire a cet endroit. On peut rediscuter de ou les mettre.}

This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables} (among others). Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}). %According to \cite{Glaeser04}, over the twentieth century, the cost of moving goods inland have declined by over 90\% in real terms.
The container was a major transport-cost reducing innovation (see \citealp{levinson2016box}). However, \cite{hummels2007} points out that trade composition effects dominate the characterization of time trends of international transport costs and that transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) have not declined much. In this paper, we investigate the robustness of this result. As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs. Rather, transport costs ``per se'' did decline, in both transport modes, and this is constitutes the main driving force behind the overall transport costs decrease, in particular for air transport.\medskip

@@

Second, we find that the main source of the downward trend in overall transport costs (that decreased by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013) does come from the reduction of transport cost ``per se'', rather than from trade composition effects. This stands in sharp contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004). Importantly, we show that the difference of results can be attributed to the new way of modeling of the additive component we offer. Allowing the additive costs to vary  across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period. Our result is more in line with common knowledge observations of the reduction of transport costs due, e.g., to the diffusion of the container.\smallskip

@@

Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data. Based on Norwegian firm-level export data in 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of ad-valorem trade costs (expressed in Norwegian crowns). Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs. Yet, our paper complements their findings in two main respects. First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of the ratio of the additive to the ad-valorem cost; by contrast, our estimation strategy enables us to uncover separately both values of the ad-valorem and the additive costs. From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect. While \citet{Irrazabal_2015} obtain that, for an import price of 100 USD, 6 USD are paid for additive trade costs, our own results point to a value of 2.7 USD paid in additive costs for maritime transport and 1.4 USD for air shipping. This comparison of results suggests that between a third and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode). \footnote{See Section \ref{sec:results_decomposition} for more details.} Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013. In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time. In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, which estimate the size and shape of per-shipment costs bases on Swiss data, and to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} and \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip

@@

The paper is built as follows. First, we estimate the values of international transport costs annually throughout the period 1974-2013 (and by transport mode), explicitly distinguishing between the additive and the ad-valorem components. This is made in Section \ref{sec:data_method}, which reports the mean values of the transport costs estimated over the period (by transport mode). Relying on these estimation results, we then provide a decomposition of the transport costs trend patterns, between what comes from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs ``per se''. Section \ref{sec:results_trends} is devoted to this analysis. Section \ref{sec:robustness} provides some additional robustness checks of our results. Section \ref{sec:conclu} concludes.

@@

Our estimation results about the size of additive transport costs can be compared to \citet{Irrazabal_2015}, which constitute the most recent evidence on this question. Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that additive costs (in Norwegian crowns) amount to 6\% of ad-valorem trade costs (expressed in Norwegian crowns), ie $\frac{t}{\tau\widetilde{p}}=6\%$ expressed in our terminology (with $\tau>1$, and picking up their estimated weighted mean value similarly to us). One important difference with respect to their work is that we can provide separate estimates for each cost component (additive and ad-valorem). From this, we can rebuilt the ratio in similar terms to them -- allowing us to gain in generality in this respect. Making use of our estimates for 2004, we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.4\% in air transport. This may sound surprisingly low, in contrast to the 6\% obtained by \citet{Irrazabal_2015}. This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case. While the database of \citet{Irrazabal_2015} allows to study trade costs in general, our database covers a subset made of the monetary international transport costs, as we start from the gap between the import and the export prices (on top of differences due to the country coverage). In this respect, our two papers can be used in a complementary fashion to infer the proportion of additive trade costs that are due to international shipment. This obeys the following reasoning. \citet{Irrazabal_2015}'s results imply that, for a import price of 100 USD, 6 USD are paid in additive trade costs. For the same import price of 100 USD, our results yield the value of 2.7 (1.4) USD paid in additive transport costs for maritime (air) transport.\footnote{In both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuilt the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$. With $x = 0.028$ and $0.014$ in vessel and air transport  in 2004, we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 0.027 and 0.014 in each transport mode respectively.} This comparison of results suggests that between 25\% (in Air) and 48\% (in Vessel) of additive trade costs are attributable to international transport. This, on top of our results summarized in Table \ref{tab:summary_results}, point out the evidence that the additive component represents a quantitative sizeable dimension of international transport costs.

\textbf{stop here. Grosse interrogation sur notre positionnement par rapport a Hummels dans la section suivante. Ai l'impresssion qu'on explique mal nos differences de resultats. On a l'air de dire que, contrairement a Hummels, les transport costs per se ont bien baisse, et que c'est ca qui compte. En regardant la figure "Characterizing the time trends: Applying Hummel's (2007) method", il me semble que c'est mal dit. Hummels ne dit pas que les couts purs n'ont pas baisse, d'ailleurs il montre qu'ils ont baisse, mais qu'au sein d'un mode de transport donne, ils sont en partie compense par les effets de composition. Nous, on montre que les couts purs ont baisse, et que les effets de composition, soit ne jouent pas (air), soit amplifient (au lieu de partiellement compenser, vessel) la baisse des couts purs. Il me semble qu'il faut reecrire l'interpretation des resultats en ce sens.}

@@

In doing so, Hummels (2007) implicitly assumes $\beta$ constant across the triplet origin country origin/product/year. In light of the economic interpretation of $\beta$, this means that the share of additive costs in total costs does not vary across these dimensions. However, our results of Section \ref{sec:results_decomposition} (on a year-to-year basis, see Appendix \ref{app:more_results}) and of Figure \ref{fig:decomp_TC_3d}, yield to cast some doubt on the empirical relevance of this assumption. As reported in Figure \ref{fig:decomp_TC_3d}, if overall transport costs have decreased over time, it is also the case for each additive and multiplicative components. In particular, the share of additive costs appears to be varying over time, in both transport modes. On top of this, one may wonder about the constancy of the share of additive costs along the couple (product/origin country).

@@

As displayed in Figure \ref{fig:histogram_beta}, for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, product, origin country) is smoothly distributed over the interval $[0,1]$, with a mode of the distribution stands around 0.3-0.4 depending of the transport mode, consistently with our previous findings. In sharp contradiction with the assumption made by  \cite{hummels_skiba} or \cite{hummels2007}, that would rather imply a distribution concentrated on a single point, this suggests that the elasticity of transport costs to the export price, or equivalently the share of additive costs in total costs, is varying across time, product and country partner.\footnote{We have checked that it is also the case when we look at the distribution over the range of origin countries or products, for a given year. These results are not reported for sake of space saving but they are available upon request to the authors.} In the next section, we shed light on the role of this result in the analysis of the  underlying sources of the overall transport costs trend patterns (between trade composition effects and transport costs ``per se''). This is made in next section.

@@

The above results point out that the share of additive costs in total transport costs is not constant across the triplet year/product/origin country, thereby contrasting the implicit Hummel's (2007) assumption. This drives us to provide a decomposition of the transport costs trends (between the trade composition effects and the ``pure'' transport costs time trends), that explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} An important contribution of the paper is to show the importance of this (empirically relevant) in analyzing the underlying determinants of transport costs trend patterns. We start presenting our estimation strategy.\smallskip

\paragraph{Estimation strategy} Estimation driven in Section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), that vary over time, product and origin country. Starting from these values, we extract the ``ceteris paribus'' transport cost measure, for each additive and multiplicative cost component (by transport mode) by the mean of a time fixed effect (see details below). For each cost component (additive and ad-valorem), comparing the unfitted measure and the fitted measure (composition effects excluded) allows to characterize if the decrease observed over the period is due to trade composition effects (for instance, changes in the country partners, in the type/ quality of products traded), or if it is the ``ceteris paribus'' costs (for instance, insurance or handling costs) that have reduced over time. We conduct the same analysis on a overall transport costs measure, built by agglomerating the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the ``unfitted'' total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported from another perspective (basis 100 in year 1974).}

The remaining question is then, how to eliminate the composition effects? Our reasoning obeys the following rationale. Starting from the values of each additive and ad-valorem transport cost measure estimated in Section \ref{sec:results_decomposition}, we extract the changes over time in the ``ceteris paribus'' transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974. One advantage of this method is to yield measures of transport costs (fitted and unfitted) that are easily comparable between transport modes and transport cost components.

\paragraph{Estimation strategy} This empirical strategy can be described as a two-stage process. First, for each transport costs component (additive/multiplicative), we decompose the estimated measure in the three product/country/time dimensions, using fixed effects.

@@

\noindent where $\mathbbm{1}_i$ and $\mathbbm{1}_{s(k)}$ represent country- and sector- fixed effects.\footnote{Throughout the exercise, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product- and year- dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows the considered year. As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, we estimate the following equation using non-linear least squares (using the same weighting scheme):\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other). Note that we impose the same weighting scheme as for the OLS regression (based on the relative value of the flow).}

@@

\noindent As a result, the two series (for $\Gamma^{adv}_t$ and $\Gamma^{add}_t$) have a straightforward interpretation in percentage changes from the initial value of 100 for $t=1974$.

Last, we rebuild a measure of total transport costs as the sum of the two components (additive and iceberg), on both the unfitted measures  and the ``ceteris paribus'' estimated transport costs (composition effects excluded). Details are reported in Appendix \ref{app:comp-effects}.

@@

\begin{figure}[htbp]
\caption{Transport costs (with and without composition effects)}
\label{fig:totalTC_compeffects_excl}
\begin{center}
\includegraphics[height=4in]
{graph_composition_all.pdf}
\end{center}
\end{figure}

Three main results emerge from Figure \ref{fig:totalTC_compeffects_excl}. First, in accordance with Figure \ref{fig:decomp_TC_3d}, we find that international transport costs have substantially decreased over the period, and in both transport modes (Panels (c) and (f)). International transport costs were reduced by 50\% between 1974 and 2013 in air shipping, and by 60\% in maritime shipping. This stands in line with the related literature (see \citealp{Hummels_1999}, \citealp{Lafourcade_Thisse}). Second, and as discussed in Section \ref{sec:results_decomposition}, the magnitude of the decrease is roughly of same order both the ad-valorem and the additive components (Panels (a) to (d)). Further, the transport cost reduction is much smoother in maritime transport, while air transport shows more volatility in the trend pattern, in particular in the 1980s and the 2000s.
Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs. Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl}, we do not find much evidence of a substantial difference between the unfitted transport costs measure and the ``ceteris paribus'' transport costs in Air transport. For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line are very close to each other, almost every year throughout the period (Figure \ref{fig:totalTC_compeffects_excl}, panels (a) to (c)). Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the ``ceteris paribus'' transport costs. Composition effects are more pronounced in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)), but once again, they do not change significantly the abovementioned decreasing trend. Considering the raw series (plain line), maritime transport costs have decreased by 60\% over the period, which can be decomposed in a 50\% decrease in transport costs \textit{per se} (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two). This is particularly the case for the multiplicative component (panel (d)).\smallskip

\textbf{VV: Figure 2 : 1) peu lisible en l’état. Soit on conserve ces graphs, et dans ce cas-la, on commente la dynamique, pas seulement le debut et la fin; si on veut rester sur seulement le debut et la fin, mettre un tableau. 2) Pour Air : lorsqu’on exclut les effets de compo, les couts multi sont plus importants, les couts additifs ne bougent pas... et les couts totaux sont en-dessous ! Déconcertant. Peut-etre est-ce du à la façon dont les couts sont repartis par secteur. 3) Pour Vessel, il y a de vrais effets de composition (sauf la dernière annee), il n’est pas possible d’affirmer qu’il n’y en a pas. Ok AI REFORMULE LES CHOSES (SAUF POUR REMARQUE 2). OK ainsi? }

@@

This decreasing time trend we find for transport costs stands in sharp contrast with \citet{hummels2007}, who obtains that the ``ceteris paribus'' transport costs had decreased more than the unfitted ones (over 1974-2004), suggesting an important role to composition effects. If anything, we find the opposite result here. This drives us to investigate further this difference of result. As we explain in more details in Appendix \ref{app:comp-effects}, our empirical strategy differs from \cite{hummels2007} in one main dimension. Our characterization of the time trends in transport costs starts from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition}), as well as for the overall transport cost (rather than the actual cif-fas price gap). As direct consequence, our methodology lets the ratio between the additive and the ad-valorem components of transport costs vary over the three time-product-partner country dimensions. %By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
This turns out to be of primary importance in the disentangling in the time trend of transport costs, between what comes from the trade composition effects and what comes from changes in the ``ceteris paribus'' transport cost dimension, as we show below. \footnote{As discussed in Appendix \ref{app:comp-effects}, our methodology also differs from  \citet{hummels2007} with respect to the weighting scheme used to obtain the evolution of the ``ceteris paribus'' transport costs over time. Robustness analysis to this point is driven in Section \ref{sec:robustness}.} To establish this point clearly, we replicate the method adopted by \cite{hummels2007} exposed above on our database (which is the same as his until 2004). The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}. The dotted line labeled ``expenditure/import value'' represents the unfitted measure of transport costs ($TC_t$ in the above terminology) and the plain line labeled ``fitted ad-valorem rate'' is the measure of transport costs composition effects excluded ($\widehat{TC}_t$).

>>>>

\begin{abstract}
This paper investigates the underlying sources behind the downward trend of international transport costs. In the same spirit as \cite{hummels2007}, we identify the respective roles of the reduction in transport costs ``per se'' and changes in trade composition effects. This drives us to provide a careful modeling of international transport costs, that allows for both ad-valorem (multiplicative) and per-unit (additive) components, exploiting information contained in the US imports flows database over 1974-2013 distinguished by transport mode, air and vessel. In contrast to \cite{hummels2007}, the main driver of the overall transport costs decrease is the reduction of transport costs ``per se''. Further, if trade composition matter, they tend to amplify (rather than reduce) the decrease in pure transport costs. Importantly, this difference of results can be attributed to the new way of modeling the additive component of transport costs we offer. Further, in quantitative terms, we find that the additive component of transport costs is sizeable, representing between 40 and 50\% of overall transport costs. In both aspects, our results point the importance of the additive component in accounting for international transport costs.
\end{abstract}

@@

This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables} (among others). Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}). %According to \cite{Glaeser04}, over the twentieth century, the cost of moving goods inland have declined by over 90\% in real terms.
The container was a major transport-cost reducing innovation (see \citealp{levinson2016box}). However, \cite{hummels2007} points out the role of trade composition effects in accounting for the time trends of international transport costs. Based on US imports over 1974-2004, he finds that transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) have decreased in both air and maritime shipment. However, he also finds that trade composition effects partially compensate this ``pure'' transport costs decrease, thereby attenuating the overall transport costs downward trend. In this paper, we investigate the robustness of this result. As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs. Rather, the marked reduction of transport costs ``per se'' constitutes the main driving force behind the overall transport costs decrease, in particular for air transport. In maritime shipment, trade composition effects matter much, but they amplify (rather than offset) the reduction in pure transport costs.\medskip

@@

Second, we find that the main source of the downward trend in overall transport costs (that decreased by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013) does come from the reduction of transport cost ``per se''. Trade composition effects matter much, and when they matter (in maritime transport), they amplify the reduction of ``pure'' transport costs, strengthening the downward trend for overall transport costs. This stands in sharp contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004). Importantly, we show that the difference of results can be attributed to the new way of modeling of the additive component we offer. Allowing the additive costs to vary  across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period. Our result thus stands in line with common knowledge observations of the reduction of transport costs due, e.g., to the diffusion of the container.\smallskip

@@

Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data. Based on Norwegian firm-level export data in 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of the ad-valorem trade costs (expressed in Norwegian crowns). Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs. Yet, our paper complements their findings in two main respects. First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of the ratio of the additive to the ad-valorem cost; by contrast, our estimation strategy enables us to uncover separately both values of the ad-valorem and the additive costs. From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect. While \citet{Irrazabal_2015} obtain that, for an import price of 100 USD, 6 USD are paid in additive trade costs, our own results point to a value of 2.7 USD paid in additive costs for maritime transport and 1.4 USD for air shipping. This comparison of results suggests that between a third and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode). \footnote{See Section \ref{sec:results_decomposition} for more details.} Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013. In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time. In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, which estimate the size and shape of per-shipment costs bases on Swiss data, and to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} and \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip

@@

The paper is built as follows. First, we estimate the values of international transport costs annually throughout the period 1974-2013 (and by transport mode), explicitly distinguishing between the additive and the ad-valorem components. This is made in Section \ref{sec:data_method}, which reports the mean values of the transport costs estimated over the period (by transport mode). Relying on these estimation results, we then provide a decomposition of the transport costs trend patterns, between what comes from changes in the trade composition (by product and/or partner country), and what is attributable to changes in transport costs ``per se''. Section \ref{sec:results_trends} is devoted to this analysis. Section \ref{sec:conclu} concludes.

@@

Our estimation results about the size of additive transport costs can be compared to \citet{Irrazabal_2015}, which constitute the most recent evidence on this question. Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that additive costs (in Norwegian crowns) amount to 6\% of ad-valorem trade costs (expressed in Norwegian crowns), ie $\frac{t}{\tau\widetilde{p}}=6\%$ expressed in our terminology (with $\tau>1$, and picking up their estimated weighted mean value similarly to us). One important difference with respect to their work is that we can provide separate estimates for each cost component (additive and ad-valorem). From this, we can rebuilt the ratio in similar terms to them -- allowing us to gain in generality in this respect. Making use of our estimates for 2004, we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.4\% in air transport. This may sound surprisingly low, in contrast to the 6\% obtained by \citet{Irrazabal_2015}. This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case. While the database of \citet{Irrazabal_2015} allows to study trade costs in general, our database covers a subset made of the monetary international transport costs, as we start from the gap between the import and the export prices (on top of differences due to the country coverage of the two databases). In this respect, our two papers can be used in a complementary fashion to infer the proportion of additive trade costs that are due to international shipment. This obeys the following reasoning. \citet{Irrazabal_2015}'s results imply that, for a import price of 100 USD, 6 USD are paid in additive trade costs. For the same import price of 100 USD, our results yield the value of 2.7 (1.4) USD paid in additive transport costs for maritime (air) transport.\footnote{In both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuilt the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$. With $x = 0.028$ and $0.014$ in vessel and air transport in 2004, we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 0.027 and 0.014 in each transport mode respectively.} This comparison of results suggests that between 25\% (in Air) and 48\% (in Vessel) of additive trade costs are attributable to international transport. On top of our results summarized in Table \ref{tab:summary_results}, this points out the evidence that the additive component represents a quantitative sizeable dimension of international transport costs. In the next section, we keep on investigating the importance of the (varying pattern of) additive costs, as regards with the determinants of transport costs trend patterns.

@@

In doing so, Hummels (2007) implicitly assumes $\beta$ constant across the triplet origin country origin/product/year. In light of the economic interpretation of $\beta$, this means that the share of additive costs in total costs does not vary across these dimensions. However, our results of Section \ref{sec:results_decomposition} (on a year-to-year basis, see Appendix \ref{app:more_results}) and of Figure \ref{fig:decomp_TC_3d}, cast some doubt on the empirical relevance of this assumption. As reported in Figure \ref{fig:decomp_TC_3d}, if overall transport costs have decreased over time, it is also the case for each additive and multiplicative components. In particular, the share of additive costs appears to be varying over time, in both transport modes. On light of this, one may wonder about the constancy of the share of additive costs along the couple (product/origin country).

@@

As displayed in Figure \ref{fig:histogram_beta}, for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, product, origin country) is smoothly distributed over the interval $[0,1]$, with a mode of the distribution stands around 0.3-0.4 depending of the transport mode, consistently with our previous findings. This stands in sharp contrast with the assumption made by  \cite{hummels_skiba} or \cite{hummels2007}, that would rather imply a distribution concentrated on a single point. Rather, this suggests that the elasticity of transport costs to the export price, or equivalently the share of additive costs in total costs, is varying across time, product and country partner.\footnote{We have checked that it is also the case when we look at the distribution over the range of origin countries or products, for a given year. These results are not reported for sake of space saving but they are available upon request to the authors.} In the next section, we shed light on the role of this result in the analysis of the underlying sources of the overall transport costs trend patterns, between trade composition effects and transport costs ``per se''.

@@

Based on the above results, we provide a decomposition of the transport costs trends between the trade composition effects and the ``pure'' transport costs time trends, that explicitly takes into account the varying share of additive costs over time, country partner and product.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} An important contribution of the paper is to show the crucial role of this (empirically relevant) assumption, as it notably modifies the conclusions relative to the role of trade composition effects. We start presenting our estimation strategy before turning to the results.\smallskip

\paragraph{Estimation strategy} Estimation driven in Section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), that vary over time, product and origin country. Starting from these values, we extract the ``ceteris paribus'' transport cost measure, for each additive and multiplicative cost component (by transport mode) by the mean of a time fixed effect. Precisely, starting from the values of each additive and ad-valorem transport cost measure estimated in Section \ref{sec:results_decomposition}, we extract the changes over time in the ``ceteris paribus'' transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974. One advantage of this method is to yield measures of transport costs (fitted and unfitted) that are easily comparable between transport modes and transport cost components.
For each cost component (additive and ad-valorem), comparing the unfitted measure and the fitted measure (composition effects excluded) allows to characterize if the decrease observed over the period is due to trade composition effects (for instance, changes in the country partners, in the type/ quality of products traded), or if it is the ``ceteris paribus'' costs (for instance, insurance or handling costs) that have reduced over time. We conduct the same analysis on a overall transport costs measure, built by agglomerating the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the ``unfitted'' total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported from another perspective (basis 100 in year 1974).}

\paragraph{Estimation method} This empirical strategy can be described as a two-stage process. First, for each transport costs component (additive/multiplicative), we decompose the estimated measure in the three product/country/time dimensions, using fixed effects.

@@

\noindent where $\mathbbm{1}_i$ and $\mathbbm{1}_{s(k)}$ represent country- and sector- fixed effects.\footnote{Throughout the exercise, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product- and year- dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows the considered year. As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, we estimate the following equation using non-linear least squares:\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other). Note that we impose the same weighting scheme as for the OLS regression (based on the relative value of the flow).}

@@

\noindent As a result, the two series $\Gamma^{adv}_t$ and $\Gamma^{add}_t$ have a straightforward interpretation in percentage changes from the initial value of 100 for $t=1974$.

Last, we rebuild a measure of total transport costs as the sum of the two components (additive and iceberg), on both the unfitted measures  and the ``ceteris paribus'' estimated transport costs (composition effects excluded). Details are reported in Appendix \ref{app:comp-effects}. \textbf{en dire plus la dessus ici? Si c'est le point majeur de la comparaison avec Hummels, il me semble que oui.}

@@

\begin{figure}[htbp]
\caption{Transport costs (with and without composition effects)}
\label{fig:totalTC_compeffects_excl}
\begin{center}
\includegraphics[height=5in]
{graph_composition_all.pdf}
\end{center}
\end{figure}

Three main results emerge from Figure \ref{fig:totalTC_compeffects_excl}. First, in accordance with Figure \ref{fig:decomp_TC_3d}, we find that international transport costs have substantially decreased over the period, and in both transport modes (Panels (c) and (f)). International transport costs were reduced by 50\% between 1974 and 2013 in air shipping, and by 60\% in maritime shipping. This stands in line with the related literature (see \citealp{Hummels_1999}, \citealp{Lafourcade_Thisse}). Second, and as discussed in Section \ref{sec:results_decomposition}, the magnitude of the decrease is roughly of same order both the ad-valorem and the additive components (Panels (a) to (d)). Further, the transport cost reduction is much smoother in maritime transport, while air transport shows more volatility in the trend pattern, in particular in the 1980s and the 2000s.
Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs. Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl}, we do not find much evidence of a substantial difference between the unfitted transport costs measure and the ``ceteris paribus'' transport costs in Air transport. For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line are very close to each other, almost every year throughout the period (Figure \ref{fig:totalTC_compeffects_excl}, panels (a) to (c)). Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the ``ceteris paribus'' transport costs. Composition effects are more pronounced in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)), where they amplify the reduction in ``pure'' transport costs (the plain line being lower than the dotted line). Considering the raw series (plain line), maritime transport costs have decreased by 60\% over the period, which can be decomposed in a 50\% decrease in transport costs \textit{per se} (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two). This is particularly the case for the multiplicative component (panel (d)).\smallskip

\textbf{VV: 2) Pour Air : lorsqu’on exclut les effets de compo, les couts multi sont plus importants, les couts additifs ne bougent pas... et les couts totaux sont en-dessous ! Déconcertant. Peut-etre est-ce du à la façon dont les couts sont repartis par secteur.}

@@

These results stand in sharp contrast with \citet{hummels2007}, who obtains that trade composition effects do matter, as they partially offset the reduction in the ``ceteris paribus'' transport costs for both air and maritime shipping. If anything, we find the opposite result here. This drives us to investigate further this difference of result. As we explain in more details in Appendix \ref{app:comp-effects}, our empirical strategy differs from \cite{hummels2007} in one main dimension. Our characterization of the time trends in transport costs starts from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition}), as well as for the overall transport cost (rather than the actual cif-fas price gap). As direct consequence, our methodology lets the ratio between the additive and the ad-valorem components of transport costs vary over the three time-product-partner country dimensions. %By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
This turns out to be of primary importance in the disentangling in the time trend of transport costs, between what comes from the trade composition effects and what comes from changes in the ``ceteris paribus'' transport cost dimension, as we show below. To establish this point clearly, we replicate the method adopted by \cite{hummels2007} exposed above on our database (which is the same as his until 2004). The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}. The dotted line labeled ``expenditure/import value'' represents the unfitted measure of transport costs ($TC_t$ in the above terminology) and the plain line labeled ``fitted ad-valorem rate'' is the measure of ``pure'' transport costs, i.e. composition effects excluded ($\widehat{TC}_t$).
